---
title: "Datenauswertung"
output: html_notebook
---

*Data Pre-Processing*

On basis of the responses to the survey, these responses were loaded from the csv-file.

```{r}
library(readr)

answers <- read_csv("Daten.csv")

```

The irrelevant columns are removed.

```{r}

answers[1:3] <- list(NULL)
answers[2:4] <- list(NULL)
answers[3:14] <- list(NULL)

```

The different characteristics of the column "Status" are displayed with their count.

```{r}

library(dplyr)

answers %>%
  group_by(Status) %>%
  summarize(count=n())

```

The second row is removed.

```{r}

answers <- answers[-c(2),]

```

The rows with the status "Umfragenvorschau" are removed because they are irrelevant.

```{r}

answers <- answers[!(answers$Status=='Umfragenvorschau'),]

```

The columns are renamed.

```{r}

newNames <- c("Status","Finished","Kauf_A1","GemüseA1_1","GemüseA1_2","GemüseA1_3",
              "GemüseA1_4","GemüseA1_5","Preis_A1","Kauf_A2","GemüseA2_1","GemüseA2_2","GemüseA2_3",
              "GemüseA2_4","GemüseA2_5","Preis_A2","Kauf_B1","GemüseB1_1","GemüseB1_2","GemüseB1_3",
              "GemüseB1_4","GemüseB1_5","Preis_B1","Kauf_B2","GemüseB2_1","GemüseB2_2","GemüseB2_3",
              "GemüseB2_4","GemüseB2_5","Preis_B2","Klima1","Klima2","Klima3","Klima4","Klima5","Klima6",
              "Alter","Geschlecht","Bildung","Einkommen","Ernährung1","Ernährung2","Beschäftigung")

names(answers) <- newNames

```

The responses to the Likert scales are transformed in numerical values.

```{r}

answers <- answers %>%
  mutate(across(c(3,10,17,24), ~ case_match(
  .,
  "Eindeutig A" ~ -2,
  "Eher A" ~ -1,
  "Ist mir egal" ~ 0,
  "Eher B" ~ 1,
  "Eindeutig B" ~ 2
)))

```

```{r}

answers <- answers %>%
  mutate(across(contains("Gemüse"),  ~ dplyr::recode(
    .,
  "Viel schlechter als A" = -2,
  "Eher schlechter als A" = -1,
  "Ungefähr gleich" = 0,
  "Eher besser als A" = 1,
  "Viel besser als A" = 2,
  .default = 1000
)))

```

The prices are transformed in the form "X.XX" and saved as a numeric datatype.

```{r}

answers[,grep("Preis",colnames(answers))] <- lapply(answers[,grep("Preis",colnames(answers))], gsub, pattern="€",replacement="")
answers[,grep("Preis",colnames(answers))] <- lapply(answers[,grep("Preis",colnames(answers))], gsub, pattern=",",replacement=".")
answers <- answers %>%
  mutate_at(grep("Preis",colnames(.)),as.numeric)

```

Also the responses to the awareness of climate change and to the eating form are transformed in numerical values.

```{r}

answers <- answers %>%
  mutate(across(contains("Klima"),  ~ dplyr::recode(
    .,
  "Stimme überhaupt nicht zu" = -2,
  "Stimme teilweise nicht zu" = -1,
  "Weder noch" = 0,
  "Stimme teilweise zu" = 1,
  "Stimme voll zu" = 2,
  .default = 1000
)))

answers <- answers %>%
  mutate(Ernährung1_kodiert = dplyr::recode(Ernährung1,"Konventionell"=1,"Vegetarisch"=2,"Vegan"=3,"Sonstiges (Bitte angeben)"=4), .after = Ernährung1)

answers <- answers %>%
  mutate(Ernährung2_kodiert=case_when(
    grepl("vegetarisch", Ernährung2)~2,
    grepl("Bio", Ernährung2)~2,
    is.na(Ernährung2)~NA,
    .default = 1),
    .after = Ernährung2)

```

To remove the unfinished responses they are filtered out.

```{r}
answers <- answers[!(answers$Finished=="Falsch" & is.na(answers$Beschäftigung) & (is.na(answers$Preis_A2) & is.na(answers$Preis_B2))),]
```

```{r}
mean(answers$Klima1, na.rm = TRUE)
```

The variables for the second vegetable of survey B is reversed because the suboptimal vegetable is here contrasting to the other displays on the left side and the Likert scale has to be reversed.

```{r}

answers <- answers %>%
  mutate(across(contains("B2"), function(x) x * -1)) 

answers$Preis_B2 <- answers$Preis_B2 * -1

```

As described above, the Likert scale has to be reversed for some statements regarding the awareness of climate change.

```{r}

answers <- answers %>%
  mutate(across(contains("Klima"), function(x) x * -1)) 

answers$Klima1 <- answers$Klima1 * -1
answers$Klima6 <- answers$Klima6 * -1

```

An unnecessary row is removed.

```{r}

answers <- answers[!answers$Status=="Antworttyp",]

```

**Demographic visualization**

The demographic variables are extracted and transformed into a table with counts and percentages.

```{r}

demoNames <- c("Alter","Geschlecht","Bildung","Einkommen","Ernährung1","Beschäftigung")

demographics <- answers[demoNames]

```

```{r}
demoStats <- setNames(data.frame(matrix(ncol = 3, nrow = 0)), c("Var1", "Freq", "Prop"))

for (i in colnames(demographics)){
  df <- as.data.frame(table(answers[i]))
  df["Prop"] <- prop.table(table(answers[i]))
  demoStats <- rbind(demoStats, df)
}

demoStats <- demoStats %>%
  rename("Category"="Var1",
         "Frequency"="Freq",
         "Percentage"="Prop")

demoStats$Percentage <- round(demoStats$Percentage * 100, digits=1)

```

```{r}
library("gt")

demoTable <- demoStats %>%
  gt() %>%
  tab_row_group(label = "Annual income",
                rows=c(15:25)) %>%
  tab_row_group(label = "Employment status",
                rows = c(30:34)) %>%
  tab_row_group(label = "Educational Background",
                rows = c(10:14)) %>%
  tab_row_group(label = "Form of nutrition",
                rows=c(26:29)) %>%
  tab_row_group(label="Gender",
                rows=c(7:9)) %>%
  tab_row_group(label = "Age",
     rows = c(1:6))

```

```{r}

demoTable <- demoTable %>%
  cols_align(
    align = "auto",
    columns = c("Category")
  ) %>%
  cols_align(
    align = "right",
    columns = c("Frequency","Percentage")
  ) %>%
  cols_width(
    c(Category) ~ px(210),
    c(Frequency) ~ px(100),
    c(Percentage) ~ px(100),
  ) %>%
  opt_row_striping()

```

```{r}

demoTable <- demoTable %>%
  tab_header(
    title = "Demographic profile of the respondents (n=175)"
  )

```

```{r}

demoTable <- demoTable %>%
  tab_options(
  table.font.name = "Times New Roman",
  table.font.color = "#000000",
  table.border.top.style = "none",
  table.border.bottom.style = "solid",
  table.border.bottom.color = "#232b2b",
  table.border.bottom.width = px(3),
  column_labels.border.top.color = "white",
  column_labels.border.top.width = px(3),
  column_labels.border.bottom.color = "#000000",
  column_labels.border.bottom.width = px(3),
  data_row.padding = px(10)
 ) %>% 
 tab_style(
  style = list(
   cell_text(
    size = px(13),
    weight = "bold",
    align = "left",
    font = "Times New Roman"
   )
  ),
  locations = list(
   cells_title(groups = "title")
  )
 ) %>% 
 tab_style(
  style = list(
   cell_text(
    size = px(11),
    align = "left"
   )
  ),
  locations = list(
   cells_title(groups = "subtitle")
  )
 ) %>% 
 tab_style(
  style = list(
   cell_text(
    size = px(11)
   ),
   cell_borders(
    sides = c("bottom", "top"),
    color = "#000000",
    weight = px(1)
   )
  ),
  locations = list(
   cells_body(gt::everything())
  )
 ) %>% 
 tab_style(
  style = list( 
   cell_text(
    size = px(11),
    weight = "bold",
    color = "#232b2b",
    font = "Times New Roman"
   )
  ),
  locations = list(
   cells_column_labels(everything())
  )
 ) %>% 
 tab_style(
  style = list( 
   cell_text(
    size = px(11),
    weight = "bold",
    color = "#232b2b",
    font = "Times New Roman"
   ),
   cell_borders(
    sides = c("bottom"),
    style = "solid",
    color = "#000000",
    weight = px(2)
   )
  ),
  locations = list(
   cells_row_groups(gt::everything())
  )
 ) %>% 
 tab_style(
  style = list( 
   cell_text(
    size = px(11),
    color = "#000000",
    font = "Times New Roman"
   ),
   cell_borders(
    sides = c("bottom", "right"),
    style = "solid",
    color = "white",
    weight = px(1)
   )
  ),
  locations = list(
   cells_stub(gt::everything()),
   cells_stubhead()
  )
 ) %>% 
 tab_style(
  style = list(
   cell_text(
    font = "Times New Roman", size = px(11), 
    color = "#000000")
  ),
  location = list(
   cells_body(columns = vars(Category))
  )
 ) 

demoTable

```

*Hypotheses testing*

Subsequently, the hypotheses are tested.

**Hypothesis 1**

For the first hypothesis the willingness to buy and to pay for suboptimal vegetables are defined in two variables.
Furthermore, Shapiro-Wilk normality tests are conducted to check the normal distribution of the data. 
The mean and the standard deviation are calculated.

```{r}

h1K <- as.vector(as.matrix(answers[,c(3,17)]))
h1Z <- as.vector(as.matrix(answers[,c(9,23)]))
shapiro.test(h1K)
shapiro.test(h1Z)
mean(h1K, na.rm = TRUE)
sd(h1K, na.rm = TRUE)
mean(h1Z, na.rm = TRUE)
sd(h1Z, na.rm = TRUE)

```

As the data isn't normally distributed, a non-parametric test in form of the Wilcoxon signed rank test has to be conducted.

```{r}

wilcox.test(h1K, alternative = "less", mu=0)
wilcox.test(h1Z, alternative = "less", mu=2.50)

```

**Hypothesis 2**

For the second hypothesis the according variables are combined to a healthiness scale and a Shapiro-Wilk normality test is conducted.

```{r}

h2 <- as.vector(as.matrix(answers[,c(4,6,8,18,20,22)]))
shapiro.test(h2)
mean(h2, na.rm = TRUE)
sd(h2, na.rm = TRUE)

```

As the data isn't normally distributed, a non-parametric test in form of the Wilcoxon signed rank test has to be conducted again.

```{r}

wilcox.test(h2, alternative = "less", mu=0)

```

**Hypothese 3**

For the third hypotheses, the averages for the perceived healthiness and for the awareness of climate change are calculated.

```{r}

answers$GemüseAvg <- rowMeans(answers[,c(4,6,8,18,20,22)], na.rm = TRUE)
answers$KlimaAvg <- rowMeans(answers[,c(31:36)],na.rm=TRUE)

```

The two variables are visualized in a scatterplot.

```{r}
library("car")
scatterplot(GemüseAvg ~ KlimaAvg, data=answers, smooth=FALSE, grid=FALSE, frame=FALSE, col = "black")

```

A linear regression is conducted.

```{r}

h3.lm <- lm(GemüseAvg ~ KlimaAvg, data = answers)
summary(h3.lm)

```

On top of the linear regression Kendall's rank correlation tau and Spearman's rank correlation rho are calculated.

```{r}

cor.test(answers$GemüseAvg, answers$KlimaAvg, method="kendall")
cor.test(answers$GemüseAvg, answers$KlimaAvg, method="spearman")

```

**Hypothesis 4**

As in hypothesis 3, this time the average willingness to buy and pay are calculated.

```{r}

answers$KaufAvg <- rowMeans(answers[,c(3,17)], na.rm = TRUE)
answers$ZahlAvg <- rowMeans(answers[,c(9,23)],na.rm=TRUE)

```

The variables are visualized in a scatterplot.

```{r}

par(mfrow=c(1,2))
scatterplot(KaufAvg ~ KlimaAvg, data=answers, smooth=FALSE, grid=FALSE, frame=FALSE, col = "black")
scatterplot(ZahlAvg ~ KlimaAvg, data=answers, smooth=FALSE, grid=FALSE, frame=FALSE, col = "black")

```

A linear regression is conducted.
First for the willingness to buy.

```{r}

h4K.lm <- lm(KaufAvg ~ KlimaAvg, data = answers)
summary(h4K.lm)

```

Afterwards for the willingness to pay.

```{r}

h4Z.lm <- lm(ZahlAvg ~ KlimaAvg, data = answers)
summary(h4Z.lm)

```

Again, the correlation coefficients are calculated.

```{r}

cor.test(answers$KaufAvg, answers$KlimaAvg, method="kendall")
cor.test(answers$KaufAvg, answers$KlimaAvg, method="spearman")
cor.test(answers$ZahlAvg, answers$KlimaAvg, method="kendall")
cor.test(answers$ZahlAvg, answers$KlimaAvg, method="spearman")

```

**Hypothese 5**

For hypothesis 5, the willingness to buy and pay have to be separated into vegetarian and vegan consumers on the one side and conventional eaters on the other side.

```{r}

answers <- answers %>%
  mutate(VegKaufAvg=if_else(Ernährung1_kodiert==2 | Ernährung1_kodiert==3 | Ernährung2_kodiert==2,
                            rowMeans(answers[,c(3,17)], na.rm = TRUE),
                            NA))

answers <- answers %>%
  mutate(VegZahlAvg=if_else(Ernährung1_kodiert==2 | Ernährung1_kodiert==3 | Ernährung2_kodiert==2,
                            rowMeans(answers[,c(9,23)], na.rm = TRUE),
                            NA))
answers <- answers %>%
  mutate(KonvKaufAvg=if_else(Ernährung1_kodiert==1 | Ernährung2_kodiert==1,
                            rowMeans(answers[,c(3,17)], na.rm = TRUE),
                            NA))

answers <- answers %>%
  mutate(KonvZahlAvg=if_else(Ernährung1_kodiert==1 | Ernährung2_kodiert==1,
                            rowMeans(answers[,c(9,23)], na.rm = TRUE),
                            NA))

```

Afterwards, Shapiro-Wilk normality tests as well as means and standard deviations are calculated.

```{r}

shapiro.test(answers$VegKaufAvg)
shapiro.test(answers$KonvKaufAvg)
shapiro.test(answers$VegZahlAvg)
shapiro.test(answers$KonvZahlAvg)

mean(answers$VegKaufAvg, na.rm = TRUE)
sd(answers$VegKaufAvg, na.rm = TRUE)
mean(answers$KonvKaufAvg, na.rm = TRUE)
sd(answers$KonvKaufAvg, na.rm = TRUE)
mean(answers$VegZahlAvg, na.rm = TRUE)
sd(answers$VegZahlAvg, na.rm = TRUE)
mean(answers$KonvZahlAvg, na.rm = TRUE)
sd(answers$KonvZahlAvg, na.rm = TRUE)

```

Wilcoxon rank sum tests have to be conducted.

```{r}

wilcox.test(answers$VegKaufAvg, answers$KonvKaufAvg, alternative="greater")
wilcox.test(answers$VegZahlAvg, answers$KonvZahlAvg, alternative="greater")

```

**Hypothese 6**

Firstly, we regard the willingness to buy. 
Shapiro-Wilk normality tests as well as means and standard deviations are calculated.

```{r}

shapiro.test(answers$Kauf_A1)
shapiro.test(answers$Kauf_A2)
mean(answers$Kauf_A1, na.rm = TRUE)
sd(answers$Kauf_A1, na.rm = TRUE)
mean(answers$Kauf_A2, na.rm = TRUE)
sd(answers$Kauf_A2, na.rm = TRUE)

```

This time a paired Wilcoxon signed rank test has to be used.

```{r}

wilcox.test(answers$Kauf_A1, answers$Kauf_A2, paired = TRUE, alternative = "less")

```

For the willingness to pay the variables have to be normalized in order to guarantee the comparability. So the indicated prices from the participants are divided by the reference prices.

```{r}

answers$Preis_A1_norm <- answers$Preis_A1 / 2.50
answers$Preis_B1_norm <- answers$Preis_B1 / 2.50
answers$Preis_A2_norm <- answers$Preis_A2 / 2.29
answers$Preis_B2_norm <- answers$Preis_B2 / 1.75

```

Shapiro-Wilk normality tests as well as means and standard deviations are calculated.

```{r}

shapiro.test(answers$Preis_A1_norm)
shapiro.test(answers$Preis_A2_norm)

mean(answers$Preis_A1_norm, na.rm = TRUE)
sd(answers$Preis_A1_norm, na.rm = TRUE)
mean(answers$Preis_A2_norm, na.rm = TRUE)
sd(answers$Preis_A2_norm, na.rm = TRUE)

```

A paired Wilcoxon signed rank test has to be used.

```{r}

wilcox.test(answers$Preis_A1_norm, answers$Preis_A2_norm, paired = TRUE, alternative = "less")

```

**Hypothesis 7**

Hypothesis 7 is tested analogous to Hypothesis 6.
Firstly, Shapiro-Wilk normality tests, means and standard deviations are calculated.

```{r}

shapiro.test(answers$Kauf_B1)
shapiro.test(answers$Kauf_B2)
shapiro.test(answers$Preis_B1_norm)
shapiro.test(answers$Preis_B2_norm)

mean(answers$Kauf_B1, na.rm = TRUE)
sd(answers$Kauf_B1, na.rm = TRUE)
mean(answers$Kauf_B2, na.rm = TRUE)
sd(answers$Kauf_B2, na.rm = TRUE)
mean(answers$Preis_B1_norm, na.rm = TRUE)
sd(answers$Preis_B1_norm, na.rm = TRUE)
mean(answers$Preis_B2_norm, na.rm = TRUE)
sd(answers$Preis_B2_norm, na.rm = TRUE)

```

Afterwards, paired Wilcoxon signed rank tests are calculated.

```{r}

wilcox.test(answers$Kauf_B1, answers$Kauf_B2, paired = TRUE, alternative = "less")
wilcox.test(answers$Preis_B1_norm, answers$Preis_B2_norm, paired = TRUE, alternative = "less")

```

**Further calculations**

***Discount***

```{r}

discP <- (2.5 - mean(answers$Preis_A1, na.rm = TRUE)) /2.5
discC <- (2.29 - mean(answers$Preis_A2, na.rm = TRUE)) /2.29
discP2 <- (2.5 - mean(answers$Preis_B1, na.rm = TRUE)) /2.5
discB <- (1.75 - mean(answers$Preis_B2, na.rm = TRUE)) /1.75
mean(c(discP,discC,discP2,discB))

```

***Health perceptions with & without label***

```{r}

healthPre <- as.vector(as.matrix(answers[,c(4,6,8,18,20,22)]))
healthPost <- as.vector(as.matrix(answers[,c(11,13,15,25,27,29)]))

shapiro.test(healthPre)
shapiro.test(healthPost)

wilcox.test(healthPre, healthPost, paired = TRUE, alternative = "less")

```

```{r}
par(mfrow=c(1,2))
plot(answers$KlimaAvg, answers$Kauf_A1, xlab="awareness of climate change", ylab="willingness to buy",
     main="without label")
abline(lm(Kauf_A1 ~ KlimaAvg, data = answers))
plot(answers$KlimaAvg, answers$Kauf_A2, xlab="awareness of climate change", ylab="willingness to buy",
     main="with label")
abline(lm(Kauf_A2 ~ KlimaAvg, data = answers))

```
***Age***

```{r}

answers %>%
  group_by(Alter) %>%
  summarize(count=n())

```

```{r}

answers <- answers %>%
  mutate(Alter_code=dplyr::recode(Alter,
  "18 - 24 Jahre" = 1,
  "25 - 34 Jahre" = 2,
  "35 - 44 Jahre" = 3,
  "45 - 54 Jahre" = 4,
  "55 - 64 Jahre" = 5,
  "65 - 74" = 6
))

```

```{r}

plot(answers$Alter_code, answers$KaufAvg, ylab="Willingness to buy", xlab = "Age", xaxt="n", main="WTB per age group")
axis(1, at=1:6, labels=c("18 - 24","25 - 34","35 - 44","45 - 54","55 - 64","65 - 74"))
abline(lm(KaufAvg ~ Alter_code, data = answers))

```

```{r}

plot(answers$Alter_code, answers$GemüseAvg, ylab="Healthiness perceptions", xlab = "Age", xaxt="n", main="Healthiness perceptions per age group")
axis(1, at=1:6, labels=c("18 - 24","25 - 34","35 - 44","45 - 54","55 - 64","65 - 74"))
abline(lm(GemüseAvg ~ Alter_code, data = answers))

```

***Gender***

```{r}

answers %>%
  group_by(Geschlecht) %>%
  summarize(count=n())

```

```{r}


answers <- answers %>%
  mutate(Geschlecht_code=dplyr::recode(Geschlecht,
  "Weiblich" = 1,
  "Männlich" = 2
))

```

```{r}

femaleWTB <- answers[answers$Geschlecht_code == 1,]$KaufAvg
femaleHP <- answers[answers$Geschlecht_code == 1,]$GemüseAvg
femaleCA <- answers[answers$Geschlecht_code == 1,]$KlimaAvg
maleWTB <- answers[answers$Geschlecht_code == 2,]$KaufAvg
maleHP <- answers[answers$Geschlecht_code == 2,]$GemüseAvg
maleCA <- answers[answers$Geschlecht_code == 2,]$KlimaAvg

```


```{r}

shapiro.test(femaleWTB)
mean(femaleWTB, na.rm = TRUE)
sd(femaleWTB, na.rm = TRUE)
shapiro.test(maleWTB)
mean(maleWTB, na.rm = TRUE)
sd(maleWTB, na.rm = TRUE)

wilcox.test(maleWTB, femaleWTB, alternative = "less")

```

```{r}

shapiro.test(femaleHP)
mean(femaleHP, na.rm = TRUE)
sd(femaleHP, na.rm = TRUE)
shapiro.test(maleHP)
mean(maleHP, na.rm = TRUE)
sd(maleHP, na.rm = TRUE)

wilcox.test(femaleHP, maleHP, alternative = "less")

```

```{r}

shapiro.test(femaleCA)
mean(femaleCA, na.rm = TRUE)
sd(femaleCA, na.rm = TRUE)
shapiro.test(maleCA)
mean(maleCA, na.rm = TRUE)
sd(maleCA, na.rm = TRUE)

wilcox.test(maleCA, femaleCA, alternative = "less")

```


